---
title: "Extreme Observation - Fitting Gamma Disribution"
author: "Subbiah"
date: "10 January 2018"
output: html_document
---
Extreme value observations are in general, asymmetric; it may be interested to study about the characteristics of tail rather than most common approach of concentrating the central part of distribution.

This notes attempts fitting Gamma distribution for a skewed data set. A quick numerical and graphical summary is helpful in understanding asymmetry and tail behaviour. Mean Excess plot may be more useful than histogram


```{r,echo=FALSE,message=FALSE, warning=FALSE}
require(moments)
require(evir)
#--------------------------------------------------------
ev_su=function(data)
  {
x1=summary(data) ; x2=var(data)
x3=skewness(data); x4=kurtosis(data)
Q1=quantile(data,0.25);Q2=quantile(data,0.50);Q3=quantile(data,0.75)
maximum=max(data)
GalSke=((Q3+Q1)-(2*Q2))/(Q3-Q1) ##galton skewness
oct7=(quantile(data,7/8));oct5=quantile(data,5/8);oct3=quantile(data,3/8)
oct1 =quantile(data,1/8);oct6=quantile(data,6/8);oct2= quantile(data,2/8)
MoorKur=((oct7-oct5+oct3-oct1)/(oct6-oct2)) ##moors kurtosis
su_re=rbind(Mean=x1[4],Median=x1[3],Variance=x2,maximum,skewness=x3,kurtosis=x4,GaltSke=GalSke,MooreKu=MoorKur)
colnames(su_re)="Summary"
su_re
}
#--------------------------------------------------------------------
#DATA
x<-c(44,	5,	17,	64,	58,	1,	52,	4,	277,	52,	7,	16,	6,	96,	51,	44,	19,	18,	83,	28,	167,	43,	12,	26,	64,	375,	14,	1,	50,	3,	40,	91,	2,	147,	65,	9,	88,	43,	24,	14,	65,	1,	53,	48,	6,	54,	21,	87,	145,	152,	20,	179,	35,	40,	74,	26,	44,	2,	1,	2,	2,	9,	78,	132,	83,	78,	14,	19,	19,	45,	103,	30,	1,	4,	1,	115,	3,	37,	15,	1,	36,	37,	55,	17,	42,	47,	93,	30,	31,	13,	89,	11,	7,	4,	39,	51,	79,	4,	33,	68,	14,	62,	3,	213,	8,	153,	100,	7,	24,	1,	67,	75,	50,	6,	5,	13,	112,	4,	2,	1,	47,	1,	4,	1,	17,	182,	39,	16,	1,	35,	28,	47,	45,	12,	1,	83,	8,	19,	91,	81,	14,	178,	40,	74,	45,	221,	130,	1,	52,	47,	55,	4,	9,	35,	28,	73,	48,	26,	110,	91,	122,	14,	42,	68,	60,	209,	10,	80,	29,	1,	191,	1,	202,	5,	72,	11,	115,	86,	34,	6,	23,	1,	1,	8,	36,	33,	400,	53,	120,	11,	44,	95,	13,	1,	7,	79,	15,	196,	4,	176,	13,	4,	130,	48,	153,	1,	5,	36,	30,	14,	13,	45,	226,	17,	5,	1,	1,	1,	83,	18,	1,	7,	120,	10,	19,	26,	11,	61,	122,	216,	1,	49)
#--------------------------------------------------------------------
round(ev_su(x),4)
```


**Associated plots are**


```{r,echo=FALSE,message=FALSE, warning=FALSE}
#plots
par(mfrow=c(1,3))
histogram=hist(x)
ME=meplot(x,main="ME Plot")
qplot(x, xi = 0, trim = NA, threshold = NA, line = TRUE,labels = TRUE, main="QQ Plot")
```



Tail distribution is studied using complementary cumulative distribution function. For testing goodness of fit, Cramer-Von-Mises (CVM) method related to extreme value observation is used beyond Kolmogorov-Smirnov (KS), Anderson-Darling (AD). 

Parameters must be estimated before assessing goodness of fit tests. Method of moments (MOM), Maximum Likelihood estimation method (MLE) and Maximization of Goodness of Fit (MGE) method are considered here.

The dataset considered for this study is extracted from the portal http://www.espncricinfo.com/ci/content/player/index.html; The game of cricket and its dynamic provides realization of extreme value observations especially in test cricket format. A batsman has utmost two innings to bat, and there is a better possibility of getting high scores. High scores like 100s, 200s or more enhance the rating and performance of players. 


```{r setup, echo=FALSE,message=FALSE, warning=FALSE}
require(fitdistrplus)
require(actuar)
require(goftest)
#-----------------------------------------------------------------------------------
ev_par_es=function(data)
{
#MLE
MLE<-fitdist(data,"gamma")
parameters=MLE$estimate
par1=parameters[1]
par2=1/parameters[2]
resmle=rbind(par1,par2)
#----------------------------------------------------------------------------------
#MOM
mom=fitdist(data,"gamma",method="mme")
parameters=mom$estimate
par3=parameters[1]
par4=1/parameters[2]
resmom=rbind(par3,par4)
#----------------------------------------------------------------------------------
#MGE_3 METHODS
MGEKS=fitdist(data,"gamma",method="mge",gof="KS")
MGEAD=fitdist(data,"gamma",method="mge",gof="AD")
MGECVM=fitdist(data,"gamma",method="mge",gof="CvM")
parameters1=MGEKS$estimate
par5=parameters1[1]
par6=1/parameters1[2]
parameters2=MGEAD$estimate
par7=parameters2[1]
par8=1/parameters2[2]
parameters3=MGECVM$estimate
par9=parameters3[1]
par10=1/parameters3[2]
resMGEKS=rbind(par5,par6);resMGEAD=rbind(par7,par8);resMGECVM=rbind(par9,par10)
#-------------------------------------------------------------------------------------
fin_res=cbind(resmom,resmle,resMGEKS,resMGEAD,resMGECVM)
row.names(fin_res)=c("Shape","Scale")
colnames(fin_res)=c("MOM","MLE","MGEKS","MGEAD","MGECVM")
fin_res
}
#--------------------------------------------------------------

#GOF--------------------------------------------------------------------------------
ev_gf=function(x)
  {
zz=ev_par_es(x)
#------------------------------------------------------------
#MLE
gofMLE_ks = ks.test(x = x, y= "pgamma", shape=zz[1,1],scale=zz[2,1])
gofMLE_AD = ad.test(x = x, null= "pgamma", shape=zz[1,1],scale=zz[2,1])
gofMLE_cvm=cvm.test(x = x, null= "pgamma", shape=zz[1,1],scale=zz[2,1])
Re_MLE1=cbind(gofMLE_ks$statistic,gofMLE_AD$statistic,gofMLE_cvm$statistic)
Re_MLE2=cbind(gofMLE_ks$p.value,gofMLE_AD$p.value,gofMLE_cvm$p.value)
Re_MLE=rbind(Re_MLE1,Re_MLE2)
#MOM
gofMOM_ks = ks.test(x = x, y= "pgamma", shape=zz[1,2],scale=zz[2,2])
gofMOM_AD = ad.test(x = x, null= "pgamma", shape=zz[1,2],scale=zz[2,2])
gofMOM_cvm=cvm.test(x = x, null= "pgamma", shape=zz[1,2],scale=zz[2,2])
Re_MOM1=cbind(gofMOM_ks$statistic,gofMOM_AD$statistic,gofMOM_cvm$statistic)
Re_MOM2=cbind(gofMOM_ks$p.value,gofMOM_AD$p.value,gofMOM_cvm$p.value)
Re_MOM=rbind(Re_MOM1,Re_MOM2)
#MGE 3 methods
gofMGEKS_ks = ks.test(x = x, y= "pgamma", shape=zz[1,3],scale=zz[2,3])
gofMGEKS_AD = ad.test(x = x, null= "pgamma", shape=zz[1,3],scale=zz[2,3])
gofMGEKS_cvm=cvm.test(x = x, null= "pgamma", shape=zz[1,3],scale=zz[2,3])
Re_MGEKS1=cbind(gofMGEKS_ks$statistic,gofMGEKS_AD$statistic,gofMGEKS_cvm$statistic)
Re_MGEKS2=cbind(gofMGEKS_ks$p.value,gofMGEKS_AD$p.value,gofMGEKS_cvm$p.value)
Re_MGEKS=rbind(Re_MGEKS1,Re_MGEKS2)

gofMGEAD_ks = ks.test(x = x, y= "pgamma", shape=zz[1,4],scale=zz[2,4])
gofMGEAD_AD = ad.test(x = x, null= "pgamma", shape=zz[1,4],scale=zz[2,4])
gofMGEAD_cvm=cvm.test(x = x, null= "pgamma", shape=zz[1,4],scale=zz[2,4])
Re_MGEAD1=cbind(gofMGEAD_ks$statistic,gofMGEAD_AD$statistic,gofMGEAD_cvm$statistic)
Re_MGEAD2=cbind(gofMGEAD_ks$p.value,gofMGEAD_AD$p.value,gofMGEAD_cvm$p.value)
Re_MGEAD=rbind(Re_MGEAD1,Re_MGEAD2)

gofMGECVM_ks = ks.test(x = x, y= "pgamma", shape=zz[1,5],scale=zz[2,5])
gofMGECVM_AD = ad.test(x = x, null= "pgamma", shape=zz[1,5],scale=zz[2,5])
gofMGECVM_cvm=cvm.test(x = x, null= "pgamma", shape=zz[1,5],scale=zz[2,5])
Re_MGECVM1=cbind(gofMGECVM_ks$statistic,gofMGECVM_AD$statistic,gofMGECVM_cvm$statistic)
Re_MGECVM2=cbind(gofMGECVM_ks$p.value,gofMGECVM_AD$p.value,gofMGECVM_cvm$p.value)
Re_MGECVM=rbind(Re_MGECVM1,Re_MGECVM2)
#----------------------
#RESULTS_GF
Fin_res_GF=cbind(Re_MOM,Re_MLE,Re_MGEKS,Re_MGEAD,Re_MGECVM)
colnames(Fin_res_GF)=c("MOMKS","MOMAD","MOMCVM","MLEKS","MLEAD","MLECVM","MGEKS-KS","MGEKS-AD","MGEKS-CVM","MGEAD-KS","MGEAD-AD","MGEAD-CVM","MGECVM-KS","MGECVM-AD","MGECVM-CVM")
rownames(Fin_res_GF)=c("Estimate","p-value")
round(Fin_res_GF,4)
}
```


**Point estimates of five approaches and results of tests of goodness of fit are provided below**
```{r, echo=FALSE,message=FALSE, warning=FALSE}
#----------------------------------------
x<-c(44,	5,	17,	64,	58,	1,	52,	4,	277,	52,	7,	16,	6,	96,	51,	44,	19,	18,	83,	28,	167,	43,	12,	26,	64,	375,	14,	1,	50,	3,	40,	91,	2,	147,	65,	9,	88,	43,	24,	14,	65,	1,	53,	48,	6,	54,	21,	87,	145,	152,	20,	179,	35,	40,	74,	26,	44,	2,	1,	2,	2,	9,	78,	132,	83,	78,	14,	19,	19,	45,	103,	30,	1,	4,	1,	115,	3,	37,	15,	1,	36,	37,	55,	17,	42,	47,	93,	30,	31,	13,	89,	11,	7,	4,	39,	51,	79,	4,	33,	68,	14,	62,	3,	213,	8,	153,	100,	7,	24,	1,	67,	75,	50,	6,	5,	13,	112,	4,	2,	1,	47,	1,	4,	1,	17,	182,	39,	16,	1,	35,	28,	47,	45,	12,	1,	83,	8,	19,	91,	81,	14,	178,	40,	74,	45,	221,	130,	1,	52,	47,	55,	4,	9,	35,	28,	73,	48,	26,	110,	91,	122,	14,	42,	68,	60,	209,	10,	80,	29,	1,	191,	1,	202,	5,	72,	11,	115,	86,	34,	6,	23,	1,	1,	8,	36,	33,	400,	53,	120,	11,	44,	95,	13,	1,	7,	79,	15,	196,	4,	176,	13,	4,	130,	48,	153,	1,	5,	36,	30,	14,	13,	45,	226,	17,	5,	1,	1,	1,	83,	18,	1,	7,	120,	10,	19,	26,	11,	61,	122,	216,	1,	49)
#-----------------------------------------------------------------------------------
par_es=ev_par_es(x)
gf_res=ev_gf(x)
par_es
gf_res
```



**Corresponding plots are**
```{r, echo=FALSE,message=FALSE, warning=FALSE}
L0=c(1:5)
L01=c("MOM","MLE","MGEKS","MGEAD","MGECVM")
par(mfrow=c(1,2))
plot(c(par_es[1,1],par_es[1,2],par_es[1,3],par_es[1,4],par_es[1,5]),xaxt="n",xlab="Shape",type = "l",ylab = "Parameter Estimation")
axis(1, at = L0, labels = L01,cex.axis=0.5)
plot(c(par_es[2,1],par_es[2,2],par_es[2,3],par_es[2,4],par_es[2,5]),xaxt="n",xlab="Scale",type = "l",ylab = "Parameter Estimation")
axis(1, at = L0, labels = L01,cex.axis=0.5)


gf_res1=gf_res[1,];gf_res2=gf_res[2,]
L1=c(1:15)
L=c("MMKS","MMAD","MMCVM","MLKS","MLAD","MLCVM","MGKSKS","MGKSAD","MGKSCVM","MGADKS","MGADAD","MGADCVM","MGCVMKS","MGCVMAD","MGCVMCVM")
par(mfrow=c(2,1))
plot(gf_res1,xaxt="n",xlab="Methods",type = "l",ylab = "Test Statistic")
axis(1, at = L1, labels = L,cex.axis=0.5)
plot(gf_res2,xaxt="n",xlab="Methods",type = "l",ylab = "p-value")
abline(h=0.05,lty=2)
axis(1, at = L1, labels = L,cex.axis=0.5)
```


#Explanation
The different behaviour of AD method can be seen from the plot. Yet, there may not be any difference as far the decision is concerned. The hypothesis of Gamma fit cannot be rejected  



